---
title: "Practical Machine Learning Project"
author: "Tim Kok"
date: "January 22, 2015"
output: html_document
---

In this project we will be predicting different types of human activity based on the measurements of wearable accelerometers. 

## Load data
The data is downloaded and loaded into a training and testing dataset. 

```{r cache=TRUE}
# create data folder
if(!file.exists('data')){dir.create('data')} 

# download training and testing data
urlTrain <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
urlTest <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
download.file(urlTrain, destfile='./data/pml-training.csv', method='curl')
download.file(urlTest, destfile='./data/pml-testing.csv', method='curl')

# load datasets (ensuring both 'NA', '#DIV/0!', and empty strings are listed NA)
training <- read.csv('data/pml-training.csv', na.strings=c("","NA", "#DIV/0!"))
testing <- read.csv('data/pml-testing.csv', na.strings=c("","NA", "#DIV/0!"))
```


## Eliminating Features
A number of features has a large number of missing values. These features are removed from the dataset, as well as certain features that contain information that is not helpful for prediction purposes (such as the name of the subject or the timestamp).
```{r}
# remove varaibles with too many NAs
sumNA <- function(x){sum(is.na(x))}
lowNAs <- lapply(training, FUN=sumNA) < 19000
training <- training[,lowNAs]
# remove column 1 to 7 with irrelevant information (name, timestamp, etc.)
training <- training[,-c(1:7)]
dim(training)
```

## Preprocessing

The initial training data will be divided into a training set and a testing set ('trainTrain', and 'testTrain'), to assess the out of sample error in cross validation. 
```{r message=FALSE}
# load caret package
library(caret)
library(ada)
# set seed for reproducibility
set.seed(4242)
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
trainTrain <- training[inTrain,] 
testTrain <- training[-inTrain,]
dim(trainTrain)
```



## 

Since we're 
```{r}
#modFit <- train(classe ~ ., data=trainTrain, method='ada')
```
To strengthen the predictors, we will use ada boost, which is recommended for classification problems. 